A Tale of Two Transects

### 2020-02-04 writing/transect_paper.Rmd
improve ASV/OTU tables

k_or_M <- function(x, ...) {
  case_when(
    abs(x) > 1e6 ~ paste(format(x / 1e6, ...), "M"),
    abs(x) > 1e3 ~ paste(format(x / 1e3, ...), "k"),
    TRUE ~ paste(format(x, ...), " ")
  )
}

@@

After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
Samples preserved using 
DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.

@@

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).'
Long (short) amplicon sequencing with PacBio yielded `r rawcounts["pb_500"]` (`r rawcounts["pb_483"]`) circular consensus reads, of which `r demuxcounts["pb_500"]` (`r demuxcounts["pb_483"]`) were successfully demultiplexed.
The 5.8S region was detected in `r regioncounts["pb_500"]` (`r regioncounts["pb_483"]`) reads.
Ion Torrent sequencing yielded `r demuxcounts["is_057"]` demultiplexed reads, with the 5.8S region detected in `r regioncounts["is_057"]`.

@@

linesep <- venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  select(starts_with("reads")) %>%
  mutate_all(nchar) %>%
  as.matrix %>%
  is_greater_than(0) %>%
  rowSums %>%
  rle() %$%
  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
  unlist() %>%
  head(length(.) - 2) %>%
  c("\\midrule")

venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-asv)",
    linesep = linesep,
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
       names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )
  )
```

@@

linesep <- venn_OTU %>%
  filter(!is.na(OTUs)) %>%
  select(starts_with("reads")) %>%
  mutate_all(nchar) %>%
  as.matrix %>%
  is_greater_than(0) %>%
  rowSums %>%
  rle() %$%
  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
  unlist() %>%
  head(length(.) - 2) %>%
  c("\\midrule")

venn_OTU %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-otu)",
    linesep = linesep,
    align = ("rcrrcrrcrr"),
    row.names = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
     " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )

>>>>

k_or_M <- function(x, ..., .sep = " ", .fun = format) {
  case_when(
    abs(x) > 1e6 ~ paste(.fun(x / 1e6, ...), "M", sep = .sep),
    abs(x) > 1e3 ~ paste(.fun(x / 1e3, ...), "k", sep = .sep),
    is.na(x) ~ NA_character_,
    TRUE ~ paste(.fun(x, ...), " ", sep = .sep)
  )
}

@@

After field lysis, DNA was extracted using the Xpedition^TM^ Soil/Fecal Prep kit (see above).
Samples preserved using LifeGuard were first centrifuged at XX ×g for Y minutes, after which the supernatant was removed and DNA was extracted from the remaining soil using the Soil/Fecal Prep kit as for the other samples.
DNA was quantified using fluorometrically using Quant-iT^TM^ PicoGreen^TM^ dsDNA (Thermo Fisher Scientific, Waltham, MA, USA) fluorescent indicator dye on a Infinite F200 plate spectrofluorometer (Tecan Trading AG, Männedorf, Switzerland) according to the manufacturer's protocol.

@@

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline are shown in Table \@ref(tab:bioinfo).
Long (short) amplicon sequencing with PacBio yielded `r rawcounts["pb_500"]` (`r rawcounts["pb_483"]`) circular consensus reads, of which `r demuxcounts["pb_500"]` (`r demuxcounts["pb_483"]`) were successfully demultiplexed.
The 5.8S region was detected in `r regioncounts["pb_500"]` (`r regioncounts["pb_483"]`) reads.
Ion Torrent sequencing yielded `r demuxcounts["is_057"]` demultiplexed reads, with the 5.8S region detected in `r regioncounts["is_057"]`.

@@

linesep <- venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  select(starts_with("reads")) %>%
  mutate_all(nchar) %>%
  as.matrix %>%
  is_greater_than(0) %>%
  rowSums %>%
  rle() %$%
  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
  unlist() %>%
  head(length(.) - 2) %>%
  c("\\midrule")
empty_background = if (options("knitr.table.format") == "html") "#777777" else "gray"
venn_ASV %>%
  filter(!is.na(ASVs)) %>%
  mutate_all(~ifelse(
    . == "",
    kableExtra::cell_spec("      ", background = empty_background, background_as_tile = FALSE), .)) %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-asv)",
    linesep = linesep,
    align = ("rcrrcrrcrr"),
    row.names = FALSE,
    escape = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
       names(venn_ASV)[-(1)] %>% 
        gsub(pattern = "(ASV_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )
  )
```

@@

linesep <- venn_OTU %>%
  filter(!is.na(OTUs)) %>%
  select(starts_with("reads")) %>%
  mutate_all(nchar) %>%
  as.matrix %>%
  is_greater_than(0) %>%
  rowSums %>%
  rle() %$%
  map(lengths, ~ c(rep("", . - 1), "\\addlinespace")) %>%
  unlist() %>%
  head(length(.) - 2) %>%
  c("\\midrule")
venn_OTU %>% as.matrix() %>% equals("") %>% which()
venn_OTU %>%
  filter(!is.na(OTUs)) %>%
  mutate_all(~ifelse(. == "", kableExtra::cell_spec("      ", background = empty_background, background_as_tile = FALSE), .)) %>%
  set_names(gsub("_[[:alpha:]]+[-._]\\d+$", "", names(.))) %>%
  set_names(gsub("_", " ", names(.))) %>%
  kable(
    booktabs = TRUE,
    caption = "(ref:venn-table-otu)",
    linesep = linesep,
    align = ("rcrrcrrcrrcrr"),
    row.names = FALSE,
    escape = FALSE
  ) %>%
  kableExtra::add_header_above(
    header = c(
      " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        rle() %$%
        set_names(lengths, plyr::mapvalues(values, datasets$seq_run, datasets$amplicon, FALSE))
      )
  ) %>%
  kableExtra::add_header_above(
    header = c(
     " " = 1,
      names(venn_OTU)[-1] %>% 
        gsub(pattern = "(OTU_)?(reads|frac)_", replacement = "") %>%
        plyr::mapvalues(datasets$seq_run, paste(datasets$tech, datasets$machine), FALSE) %>%
        rle() %$%
        set_names(lengths, values)
      )
  ) %>%
  kableExtra::landscape()
```
